Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-07-03 manuscript/pnas_style/manuscript.tex
minor organization changes

\significancestatement{
A long-standing problem in evolutionary biology is to understand the processes
that shape the genetic composition of populations. In a population without
migration, the two processes that change allele frequencies are selection,
which increases beneficial alleles and removes deleterious ones, and genetic
drift which randomly changes frequencies as some parents contribute more or
less alleles to the next generation. Previous efforts to disentangle these
processes have used genomic samples from a single timepoint and models of how
selection affects neighboring sites (linked selection). Here, we use genomic
data taken through time to quantify the contributions of selection and drift to
genome-wide frequency changes. We show selection acts over short timescales in
three evolve-and-resequence studies and has a sizeable genome-wide impact.}

@@

\dropcap{A} long-standing problem in evolutionary genetics is quantifying the
roles of genetic drift and selection in shaping genome-wide allele frequency
changes.  Selection can affect allele frequencies, both directly and
indirectly, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome, e.g. linked selection
\cite{Maynard_Smith1974-zr,Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review. Previous work has mostly focused on teasing
apart the impacts of drift and selection on genome-wide diversity using
population samples from a single contemporary timepoint, often by modeling the
correlation between regional recombination rate, gene density, and diversity
created in the presence of linked selection \cite{Cutter2013-ba,Sella2009-nx}.
This approach has shown linked selection has a major role in shaping patterns
of genome-wide diversity across the genomes of a range of sexual species
\cite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\citep{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
However, we lack an understanding of both how linked selection acts over short
time intervals and \vb{of} its full impact on genome-wide allele frequency changes.

There are numerous examples of rapid phenotypic adaptation
\cite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\cite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature of
fitness makes detecting the impact of selection on genome-wide variation over
short timescales in sexual populations remarkably difficult \cite{Latta1998-me,
Pritchard2010-tk,Kemper2014-bx}. This is because the effect of selection on a
polygenic trait (such as fitness) is distributed across numerous loci. This can
lead to subtle allele frequency shifts on standing variation that are difficult
to distinguish from background levels of genetic drift and sampling variance.
Increasingly, genomic experimental evolution studies with multiple timepoints,
and in some cases multiple replicate populations, are being used to detect
large-effect selected loci \cite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \cite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
addition these temporal-genomic studies have begun in wild populations, some
with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \cite{Bergland2014-ij,Machado2018-cs}. In a previous
paper, we proposed that one useful signal for understanding the genome-wide
impact of polygenic linked selection detectable from temporal genomic data is
the temporal autocovariance (i.e. covariance between two timepoints) \vb{of}
allele frequency changes \cite{Buffalo2019-io}. These covariances are created
when the loci that underly heritable fitness variation perturb the frequencies
of linked alleles; in contrast, when genetic drift acts alone in a closed
population, these covariances are expected to be zero for neutral alleles.
Mathematically, temporal covariances are useful because it is natural to
decompose the total variance in allele frequency change across a time interval
into the variances and covariances in allele frequency change between
generations.  Furthermore, biologically, these covariances reflect the extent
to which allele frequency changes in one generation predict changes in another
due to a shared selection pressures and associations to selected loci.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \cite{Barghi2019-qy,Kelly2019-dc}, and an artificial
selection experiment \cite{Castro2019-uk}. These sequencing selection
experiments have started to uncover selected loci contributing to the adaptive
response; however it is as yet far from clear how much of \vb{genome-wide}
allele frequency \vb{changes} are driven by selection or genetic drift. We
repeatedly find a signal of temporal covariance, consistent with linked
selection acting to significantly perturb genome-wide allele frequency changes
across the genome in a manner that other approaches would not be able
differentiate from genetic drift. We estimate a lower bound of the fraction of
variance in allele frequency change caused by selection, as well as the
correlation between allele frequency changes between replicate populations
caused by convergent selection pressures. Overall, we demonstrate that linked
selection has a powerful role in shaping genome-wide allele frequency changes
over very short timescales \vb{in experimental evolution}.

@@

We first analyzed \vb{the dataset of} Barghi et al. (2019), an
evolve-and-resequence study with ten replicate populations exposed to a high
temperature lab environment and evolved for 60 generations, and sequenced every
ten generations. Using the seven timepoints and ten replicate populations, we
estimated the genome-wide $6 \times 6$ temporal covariance matrix $\mathbf{Q}$
for each of the ten replicates. Each row of these matrices represent the
temporal covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the
allele frequency change (in ten-generation intervals, denoted $\Delta_{_{10}}
p_t$) of some initial reference generation $s$ (the row of the matrix), and
some later timepoint $t$ (the column of the matrix). We corrected these
matrices for biases created due to sampling noise, and normalized the entries
for heterozygosity (see SI Appendix, sections \ref{supp:ind-depth-var-corr} and
\ref{supp:matrix-correction}). These covariances are expected to be zero when
only drift is acting, as only heritable variation for fitness can create
covariance between allele frequency changes in a closed population
\cite{Buffalo2019-io}.  Averaging across the ten replicate temporal covariances
matrices, we find temporal covariances that are statistically significant (95\%
block bootstraps CIs do not contain zero), consistent with linked selection
perturbing genome-wide allele frequency changes over very short time periods.
The covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals (Figure
\ref{fig:figure-1} A), as expected when directional selection affects linked
variants' frequency trajectories until ultimately linkage disequilibrium and
the associated additive genetic variance for fitness decays (which could occur
as a population reaches a new optimum, \vb{and directional selection weakens})
\cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
this general pattern holds; see SI Appendix, Fig.
\ref{suppfig:barghi-cov-panels}.

Since our covariances are averages over loci, the covariance estimate could be
strongly affected by a few outlier regions. To test whether large outlier
regions drive the genome-wide signal we see in the Barghi et al. (2019) data,
we calculate the covariances in 100kb windows along the genome (we refer to
these as windowed covariances throughout) and take the median windowed
covariance, and trimmed-mean windowed covariance, as a measure of the
genome-wide covariance robust to large-effect loci. These robust estimates (SI
Appendix, Table \ref{supp:table-trimmed-mean} and SI Appendix, Fig.
\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
covariance, \vb{establishing} that genomic temporal covariances are non-zero
due to the impact of selection acting across many genomic regions.

@@

  \caption{{\bf A}: Temporal covariance, averaged across all ten replicate
    populations, through time from the Barghi et al. (2019) study. Each line
    depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from some
    reference generation $s$ to a later time $t$ which varies along the x-axis;
    each line corresponds to a row of the upper-triangle of the temporal
    covariance matrix with the same color (upper right). The ranges around each
    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: \vb{A lower
    bound on the} proportion of the total variance in allele frequency change
    explained by linked selection, $G(t)$, as it varies through time $t$ along
    the x-axis. The black line is the $G(t)$ averaged across replicates, with
    the $95\%$ block-bootstrap confidence interval. The other lines are the
    $G(t)$ for each individual replicate, with colors indicating what subset of
    the temporal-covariance matrix to the right is being included in the
    calculation of $G(t)$.}

@@

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure is not easily
interpretable. We can calculate a more intuitive measure from the temporal
covariances to quantify the impact of selection on allele frequency change: the
ratio of total covariance in allele frequency change to the total variance in
allele frequency change. We denote the change in allele frequency as $\Delta
p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation $t$.
Since the total variation in allele frequency change can be partitioned into
variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
\var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
\Delta p_j)$ (we \vb{correct for biases due to} sequencing depth), and the
covariances are zero when drift acts alone, this is a lower bound on how much
of the variance in allele frequency change is caused by linked selection
\cite{Buffalo2019-io}. We call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}.
\end{align}
%
\vb{This estimates the impact of selection on allele frequency change between
the initial generation $0$ and some later generation $t$, which can be varied
to see how this quantity grows through time}. When the sum of the covariances
is positive, this measure can intuitively be understood as \vb{a lower bound on
} relative fraction of allele frequency change normally thought of as ``drift''
that is actually due to selection. Additionally, $G(t)$ can be understood as a
short-timescale estimate of the reduction in neutral diversity due to linked
selection (or equivalently the reduction in neutral effective population size
needed to account linked selection, see SI Appendix, section
\ref{supp:reduction-approx}). Since the Barghi et al. (2019) experiment is
sequenced every ten generations, the numerator uses the covariances estimated
between ten-generation blocks of allele frequency change; thus the strong,
unobservable, covariances between adjacent generations do not contribute to the
numerator of $G(t)$. Had these covariances been measurable on shorter
timescales, their cumulative effect would likely have been higher yet (see SI
Appendix, sections \ref{supp:barghi-covs} and \ref{supp:tempblocks} for more
details).  \vb{Additionally, selection inflates the variance in allele
frequency change per generation}; however, this effect cannot be easily
distinguished from drift. For both these reasons, our measure $G(t)$ is quite
conservative (we demonstrate this through simulations in SI Appendix, section
\ref{supp:tempblocks}).  Still, we find a remarkably strong signal. Greater
than $20\%$ of total, genome-wide allele frequency change over 60 generations
is the result of selection (Figure \ref{fig:figure-1} B). This proportion of
variance attributable to selection builds over time in Figure
\ref{fig:figure-1}B as the effects of linked selection are compounded over the
generations unlike genetic drift. Our G(t) starts to plateau to a constant
level as the covariances from earlier generations have decayed and so no longer
contribute as strongly (Figure \ref{fig:figure-1}).

@@

The replicate design of Barghi et al. (2019) allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: alleles linked to a particular fitness background are expected to
have allele frequency changes in the same direction if the selection pressures
are similar. Intuitively, where temporal covariances reflect that alleles
associated with heritable fitness backgrounds are predictive of frequency
changes between generations, replicate covariances reflect that heritable
fitness backgrounds common to each replicate predict (under the same selection
pressures) frequency changes between replicates; \vb{we note that there is not
  a direct one-to-one correspondence between temporal and replicate
  covariances, since the latter are driven by a shared selection pressure and
  the stochastic genetic backgrounds across replicate populations}. We measure
  this through a statistic similar to a correlation, which we call the
  convergent correlation: the ratio of average between-replicate covariance
  across all pairs to the average standard deviation across all pairs of
  replicates,

@@

  \caption{Forward-in-time simulations demonstrate how temporal covariance,
    $G(t)$ trajectories, and convergence correlations arise during optima
    shifts of two different magnitudes, under Gaussian stabilizing selection.
    ({\bf A}) Trait means across 30 replicate before and after optima shifts
    (solid lines), for two different magnitudes (indicated by color). The
    optimal trait values are indicated by the purple and yellow dashed lines.
    ({\bf B}) Mean temporal covariance $\cov(\Delta p_5, \Delta p_t)$ across 30
    simulation replicates, where $t$ varies along the x-axis (points), with a
    loess-smoothed average (solid line). ({\bf C}) $G(t)$ trajectories through
    time, for 30 replicate simulations across two optima shifts. The solid line
    is a loess-smoothed average. ({\bf D}) The convergence correlations between
    two populations (each $1000$ diploids) split from a common population, that
    underwent either an optima shift in the same direction (converge) and
    opposite directions (diverge) at generation five. In subfigures ({\bf B}),
    ({\bf C}), and ({\bf D}), directional selection begins at generation five,
    when the optima shifts; this is indicated by the vertical dashed red line
    \vb{(see SI Appendix Section \ref{supp:tempcov-gss} for details on these
    simulations).}}

@@

\vb{Averaging across replicates, these simulation results show $G(t)$ is
  relatively insensitive to the number of loci underlying the trait. However,
  if only a small number of loci influence the trait, the $G(t)$ trajectories
  are typically much more stochastic \emph{across} replicates. This reaffirms
that the genome-wide linked selection response we see in the Barghi et al.
(2019) data is highly polygenic (compare Figure \ref{fig:figure-1}B to SI
Appendix, Fig.  \ref{suppfig:sim-expfit-Gs})}. \vb{Furthermore,} using our
simulations we find that sampling only every 10 generations does indeed mean
that our estimates of $G(t)$ are an underestimate of the \vb{proportional
effect} of linked selection as they cannot include the covariance
between closely spaced generations (see SI Appendix, Fig.
\ref{suppfig:supp-blocks}).

Additionally, we explored other \vb{modes of} selection \vb{with simulations}.
We find that the long term dynamics of the covariances under directional
truncation selection, which generates substantial epistasis, are richer than we
see under Gaussian Stabilizing Selection (GSS) and multiplicative selection (SI
Appendix, Fig.  \ref{suppfig:supp-trunc}). We also conducted simulations of
purifying selection alone (i.e. background selection) and find that this can
also generate positive temporal covariances (SI Appendix, Fig.
\ref{suppfig:supp-bgs-covs}) and under some circumstances, can even generate
convergence correlations (SI Appendix, Fig.  \ref{suppfig:supp-bgs-converg}).
Thus it is unlikely that the signatures of linked selection we see are entirely
the result of the novel selection pressure the populations are exposed to, and
some of this \vb{signature} may be ongoing purifying selection. Only in the
case of the Longshanks experiment, does the control line allow us to conclude
that selection that is \vb{almost entirely} due to the novel selection
pressure.

While none of our experiments have selected the populations in divergent
directions, in our simulations we find that such selection can generate
negative convergent correlations (Figure \ref{fig:figure-4}D). This suggests
that selection experiments combining multiple replicates, control lines, as
well as divergent selection pressures might be quite informative \vb{in
disentangling} the contribution of particular selection pressures \vb{from}
genome-wide allele frequency changes.

@@

Since the seminal analysis of Maynard Smith and Haig (1974) demonstrating that
linked neutral diversity is reduced as an advantageous polymorphism sweeps to
fixation, over four decades of theoretical and empirical research has bettered
our understanding of linked selection.  One under-used approach to understand
the genome-wide effects of selection on polygenic trait (e.g.  on standing
variation, stems from an early quantitative genetic model of linked selection
\cite{Robertson1961-ho} and its later developments
\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
models). Implicit in these models is that autocovariance between allele
frequency change is created when there is heritable fitness variation in the
population, a signal that may be readily detected from temporal genomic data
\cite{Buffalo2019-io}.  Depending on how many loci affect fitness, even a
strong effect of linked selection may not be differentiable from genetic drift
using only single contemporary population samples or looking at temporal allele
frequency change at each locus in isolation. In this way, averaging summaries
of temporal data allows us to sidestep the key problem of detecting selection
from standing variation: that the genomic footprint leaves too soft of a
signature to differentiate from a background of genetic drift. In fact we find
that the temporal covariance signal is detectable even in the extremely
\vb{difficult to detect case of selection on highly polygenic traits
\cite{Buffalo2019-io}}.

It is worth building some intuition why temporal covariance allows us to detect
such faint signals of polygenic linked selection from temporal genomic data.
\vb{Variance} in allele frequency change \vb{is subject to both} drift and
sampling noise, which at any \vb{single} locus may swamp the temporal
covariance signal \vb{due to selection, or} create \vb{spurious} covariances
\vb{when selection is not acting}.  However, these spurious covariances do not
share a directional signal whereas the covariances created by linked selection
do; consequently, averaging across the entire genome, the temporal signal
exceeds sampling noise.

>>>>

\significancestatement{
A long-standing problem in evolutionary biology is to understand the processes
that shape the genetic composition of populations. In a population without
migration, the two processes that change allele frequencies are selection,
which increases beneficial alleles and removes deleterious ones, and genetic
drift which randomly changes frequencies as some parents contribute more or
less alleles to the next generation. Previous efforts to disentangle these
processes have used genomic samples from a single timepoint and models of how
selection affects neighboring sites (linked selection). Here, we use genomic
data taken through time to quantify the contributions of selection and drift to
genome-wide frequency changes. We show selection acts over short timescales in
three evolve-and-resequence studies and has a sizable genome-wide impact.}

@@

\dropcap{A} long-standing problem in evolutionary genetics is quantifying the
roles of genetic drift and selection in shaping genome-wide allele frequency
changes.  Selection can affect allele frequencies, both directly and
indirectly, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome, e.g. linked selection
\cite{Maynard_Smith1974-zr,Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review. Previous work has mostly focused on teasing
apart the impacts of drift and selection on genome-wide diversity using
population samples from a single contemporary timepoint, often by modeling the
correlation between regional recombination rate, gene density, and diversity
created in the presence of linked selection \cite{Cutter2013-ba,Sella2009-nx}.
This approach has shown linked selection has a major role in shaping patterns
of genome-wide diversity across the genomes of a range of sexual species
\cite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\citep{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
However, we lack an understanding of both how linked selection acts over short
time intervals and of its full impact on genome-wide allele frequency changes.

There are numerous examples of rapid phenotypic adaptation
\cite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\cite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature of
fitness makes detecting the impact of selection on genome-wide variation over
short timescales in sexual populations remarkably difficult \cite{Latta1998-me,
Pritchard2010-tk,Kemper2014-bx}. This is because the effect of selection on a
polygenic trait (such as fitness) is distributed across numerous loci. This can
lead to subtle allele frequency shifts on standing variation that are difficult
to distinguish from background levels of genetic drift and sampling variance.
Increasingly, genomic experimental evolution studies with multiple timepoints,
and in some cases multiple replicate populations, are being used to detect
large-effect selected loci \cite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \cite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
addition these temporal-genomic studies have begun in wild populations, some
with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \cite{Bergland2014-ij,Machado2018-cs}. In a previous
paper, we proposed that one useful signal for understanding the genome-wide
impact of polygenic linked selection detectable from temporal genomic data is
the temporal autocovariance (i.e. covariance between two timepoints) of
allele frequency changes \cite{Buffalo2019-io}. These covariances are created
when the loci that underly heritable fitness variation perturb the frequencies
of linked alleles; in contrast, when genetic drift acts alone in a closed
population, these covariances are expected to be zero for neutral alleles.
Mathematically, temporal covariances are useful because it is natural to
decompose the total variance in allele frequency change across a time interval
into the variances and covariances in allele frequency change between
generations.  Furthermore, biologically, these covariances reflect the extent
to which allele frequency changes in one generation predict changes in another
due to a shared selection pressures and associations to selected loci.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \cite{Barghi2019-qy,Kelly2019-dc}, and an artificial
selection experiment \cite{Castro2019-uk}. These sequencing selection
experiments have started to uncover selected loci contributing to the adaptive
response; however it is as yet far from clear how much of genome-wide
allele frequency changes are driven by selection or genetic drift. We
repeatedly find a signal of temporal covariance, consistent with linked
selection acting to significantly perturb genome-wide allele frequency changes
across the genome in a manner that other approaches would not be able
differentiate from genetic drift. We estimate a lower bound of the fraction of
variance in allele frequency change caused by selection, as well as the
correlation between allele frequency changes between replicate populations
caused by convergent selection pressures. Overall, we demonstrate that linked
selection has a powerful role in shaping genome-wide allele frequency changes
over very short timescales in experimental evolution.

@@

We first analyzed the dataset of Barghi et al. (2019), an
evolve-and-resequence study with ten replicate populations exposed to a high
temperature lab environment and evolved for 60 generations, and sequenced every
ten generations. Using the seven timepoints and ten replicate populations, we
estimated the genome-wide $6 \times 6$ temporal covariance matrix $\mathbf{Q}$
for each of the ten replicates. Each row of these matrices represent the
temporal covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the
allele frequency change (in ten-generation intervals, denoted $\Delta_{_{10}}
p_t$) of some initial reference generation $s$ (the row of the matrix), and
some later timepoint $t$ (the column of the matrix). We corrected these
matrices for biases created due to sampling noise, and normalized the entries
for heterozygosity (see SI Appendix, sections \ref{supp:ind-depth-var-corr} and
\ref{supp:matrix-correction}). These covariances are expected to be zero when
only drift is acting, as only heritable variation for fitness can create
covariance between allele frequency changes in a closed population
\cite{Buffalo2019-io}.  Averaging across the ten replicate temporal covariances
matrices, we find temporal covariances that are statistically significant (95\%
block bootstraps CIs do not contain zero), consistent with linked selection
perturbing genome-wide allele frequency changes over very short time periods.
The covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals (Figure
\ref{fig:figure-1} A), as expected when directional selection affects linked
variants' frequency trajectories until ultimately linkage disequilibrium and
the associated additive genetic variance for fitness decays (which could occur
as a population reaches a new optimum, and directional selection weakens)
\cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
this general pattern holds; see SI Appendix, Fig.
\ref{suppfig:barghi-cov-panels}.

Since our covariances are averages over loci, the covariance estimate could be
strongly affected by a few outlier regions. To test whether large outlier
regions drive the genome-wide signal we see in the Barghi et al. (2019) data,
we calculate the covariances in 100kb windows along the genome (we refer to
these as windowed covariances throughout) and take the median windowed
covariance, and trimmed-mean windowed covariance, as a measure of the
genome-wide covariance robust to large-effect loci. These robust estimates (SI
Appendix, Table \ref{supp:table-trimmed-mean} and SI Appendix, Fig.
\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
covariance, establishing that genomic temporal covariances are non-zero
due to the impact of selection acting across many genomic regions.

@@

  \caption{{\bf A}: Temporal covariance, averaged across all ten replicate
    populations, through time from the Barghi et al. (2019) study. Each line
    depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from some
    reference generation $s$ to a later time $t$ which varies along the x-axis;
    each line corresponds to a row of the upper-triangle of the temporal
    covariance matrix with the same color (upper right). The ranges around each
    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: A lower
    bound on the proportion of the total variance in allele frequency change
    explained by linked selection, $G(t)$, as it varies through time $t$ along
    the x-axis. The black line is the $G(t)$ averaged across replicates, with
    the $95\%$ block-bootstrap confidence interval. The other lines are the
    $G(t)$ for each individual replicate, with colors indicating what subset of
    the temporal-covariance matrix to the right is being included in the
    calculation of $G(t)$.}

@@

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure is not easily
interpretable. We can calculate a more intuitive measure from the temporal
covariances to quantify the impact of selection on allele frequency change: the
ratio of total covariance in allele frequency change to the total variance in
allele frequency change. We denote the change in allele frequency as $\Delta
p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation $t$.
Since the total variation in allele frequency change can be partitioned into
variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
\var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
\Delta p_j)$ (we correct for biases due to sequencing depth), and the
covariances are zero when drift acts alone, this is a lower bound on how much
of the variance in allele frequency change is caused by linked selection
\cite{Buffalo2019-io}. We call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}.
\end{align}
%
This estimates the impact of selection on allele frequency change between
the initial generation $0$ and some later generation $t$, which can be varied
to see how this quantity grows through time. When the sum of the covariances
is positive, this measure can intuitively be understood as a lower bound on
relative fraction of allele frequency change normally thought of as ``drift''
that is actually due to selection. Additionally, $G(t)$ can be understood as a
short-timescale estimate of the reduction in neutral diversity due to linked
selection (or equivalently the reduction in neutral effective population size
needed to account linked selection, see SI Appendix, section
\ref{supp:reduction-approx}). Since the Barghi et al. (2019) experiment is
sequenced every ten generations, the numerator uses the covariances estimated
between ten-generation blocks of allele frequency change; thus the strong,
unobservable, covariances between adjacent generations do not contribute to the
numerator of $G(t)$. Had these covariances been measurable on shorter
timescales, their cumulative effect would likely have been higher yet (see SI
Appendix, sections \ref{supp:barghi-covs} and \ref{supp:tempblocks} for more
details).  Additionally, selection inflates the variance in allele
frequency change per generation; however, this effect cannot be easily
distinguished from drift. For both these reasons, our measure $G(t)$ is quite
conservative (we demonstrate this through simulations in SI Appendix, section
\ref{supp:tempblocks}).  Still, we find a remarkably strong signal. Greater
than $20\%$ of total, genome-wide allele frequency change over 60 generations
is the result of selection (Figure \ref{fig:figure-1} B). This proportion of
variance attributable to selection builds over time in Figure
\ref{fig:figure-1}B as the effects of linked selection are compounded over the
generations unlike genetic drift. Our G(t) starts to plateau to a constant
level as the covariances from earlier generations have decayed and so no longer
contribute as strongly (Figure \ref{fig:figure-1}).

@@

The replicate design of Barghi et al. (2019) allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: alleles linked to a particular fitness background are expected to
have allele frequency changes in the same direction if the selection pressures
are similar. Intuitively, where temporal covariances reflect that alleles
associated with heritable fitness backgrounds are predictive of frequency
changes between generations, replicate covariances reflect that heritable
fitness backgrounds common to each replicate predict (under the same selection
pressures) frequency changes between replicates; we note that there is not
  a direct one-to-one correspondence between temporal and replicate
  covariances, since the latter are driven by a shared selection pressure and
  the stochastic genetic backgrounds across replicate populations. We measure
  this through a statistic similar to a correlation, which we call the
  convergent correlation: the ratio of average between-replicate covariance
  across all pairs to the average standard deviation across all pairs of
  replicates,

@@

  \caption{Forward-in-time simulations demonstrate how temporal covariance,
    $G(t)$ trajectories, and convergence correlations arise during optima
    shifts of two different magnitudes, under Gaussian stabilizing selection.
    ({\bf A}) Trait means across 30 replicate before and after optima shifts
    (solid lines), for two different magnitudes (indicated by color). The
    optimal trait values are indicated by the purple and yellow dashed lines.
    ({\bf B}) Mean temporal covariance $\cov(\Delta p_5, \Delta p_t)$ across 30
    simulation replicates, where $t$ varies along the x-axis (points), with a
    loess-smoothed average (solid line). ({\bf C}) $G(t)$ trajectories through
    time, for 30 replicate simulations across two optima shifts. The solid line
    is a loess-smoothed average. ({\bf D}) The convergence correlations between
    two populations (each $1000$ diploids) split from a common population, that
    underwent either an optima shift in the same direction (converge) and
    opposite directions (diverge) at generation five. In subfigures ({\bf B}),
    ({\bf C}), and ({\bf D}), directional selection begins at generation five,
    when the optima shifts; this is indicated by the vertical dashed red line
    (see SI Appendix Section \ref{supp:tempcov-gss} for details on these
    simulations).}

@@

Averaging across replicates, these simulation results show $G(t)$ is relatively
insensitive to the number of loci underlying the trait. However, if only a
small number of loci influence the trait, the $G(t)$ trajectories are typically
much more stochastic \emph{across} replicates. This reaffirms that the
genome-wide linked selection response we see in the Barghi et al.  (2019) data
is highly polygenic (compare Figure \ref{fig:figure-1}B to SI Appendix, Fig.
\ref{suppfig:sim-expfit-Gs}). Furthermore, using our simulations we find
that sampling only every 10 generations does indeed mean that our estimates of
$G(t)$ are an underestimate of the proportional effect of linked selection
as they cannot include the covariance between closely spaced generations (see
SI Appendix, Fig.  \ref{suppfig:supp-blocks}).

Additionally, we explored other modes of selection with simulations.
We find that the long term dynamics of the covariances under directional
truncation selection, which generates substantial epistasis, are richer than we
see under Gaussian Stabilizing Selection (GSS) and multiplicative selection (SI
Appendix, Fig.  \ref{suppfig:supp-trunc}). We also conducted simulations of
purifying selection alone (i.e. background selection) and find that this can
also generate positive temporal covariances (SI Appendix, Fig.
\ref{suppfig:supp-bgs-covs}) and under some circumstances, can even generate
convergence correlations (SI Appendix, Fig.  \ref{suppfig:supp-bgs-converg}).
Thus it is unlikely that the signatures of linked selection we see are entirely
the result of the novel selection pressure the populations are exposed to, and
some of this signature may be ongoing purifying selection. Only in the
case of the Longshanks experiment, does the control line allow us to conclude
that selection that is almost entirely due to the novel selection
pressure.

While none of our experiments have selected the populations in divergent
directions, in our simulations we find that such selection can generate
negative convergent correlations (Figure \ref{fig:figure-4}D). This suggests
that selection experiments combining multiple replicates, control lines, as
well as divergent selection pressures might be quite informative in
disentangling the contribution of particular selection pressures from
genome-wide allele frequency changes.

@@

Since the seminal analysis of Maynard Smith and Haig (1974) demonstrating that
linked neutral diversity is reduced as an advantageous polymorphism sweeps to
fixation, over four decades of theoretical and empirical research has bettered
our understanding of linked selection.  One under-used approach to understand
the genome-wide effects of selection on polygenic trait (e.g.  on standing
variation, stems from an early quantitative genetic model of linked selection
\cite{Robertson1961-ho} and its later developments
\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
models). Implicit in these models is that autocovariance between allele
frequency change is created when there is heritable fitness variation in the
population, a signal that may be readily detected from temporal genomic data
\cite{Buffalo2019-io}.  Depending on how many loci affect fitness, even a
strong effect of linked selection may not be differentiable from genetic drift
using only single contemporary population samples or looking at temporal allele
frequency change at each locus in isolation. In this way, averaging summaries
of temporal data allows us to sidestep the key problem of detecting selection
from standing variation: that the genomic footprint leaves too soft of a
signature to differentiate from a background of genetic drift. In fact we find
that the temporal covariance signal is detectable even in the extremely
difficult to detect case of selection on highly polygenic traits
\cite{Buffalo2019-io}.

It is worth building some intuition why temporal covariance allows us to detect
such faint signals of polygenic linked selection from temporal genomic data.
Variance in allele frequency change is subject to both drift and
sampling noise, which at any single locus may swamp the temporal
covariance signal due to selection, or create spurious covariances
when selection is not acting.  However, these spurious covariances do not
share a directional signal whereas the covariances created by linked selection
do; consequently, averaging across the entire genome, the temporal signal
exceeds sampling noise.

Our analyses reveal that a sizable proportion of allele frequency change in
these experimental evolution populations is due to the (likely indirect) action
of selection. Capitalizing on replicated designs, we characterized the extent
to which convergent selection pressures lead to parallel changes in allele
frequencies across replicate populations, and found that a substantial
proportion of the response is shared across short timescales.  These likely
represent substantial under-estimates of the contribution of linked selection
because the studies we have reanalyzed do not sequence the population each
generation, preventing us from including the effects of stronger correlations
between adjacent generations. Furthermore, our estimation methods are
intentionally conservative, for example they exclude the contribution of
selection that does not persist across generations and selection that reverses
sign; thus they can be seen as a lower bound of the effects of selection, which
we have confirmed through forward-in-time simulations.  Finally, through
simulation results, we show that for a given level of additive genetic
variance, the strengths of temporal and replicate covariances depend on the
mode of selection, the details of the populations or selection experiment, and
the level of linkage disequilibria, yet the level of temporal covariance is
relatively invariant to the number of loci underlying fitness, as long as
fitness is sufficiently polygenic.
